Structured Versus Decorated Cospans

### 2022-03-15 ResponseToReviews.tex

{\chris Indeed this comment refers to the functor $\mathsf{Rex} \to \mathsf{SMC}$, I also chose to disregard R for the same reasons. The paragraph 
below existed in part later in text, and was in part expanded according to comment 29, not this one! I think I agree with that comment in that the 
assumptions of the theorem deserve a bit of explanation, and tried my best there but possibly did not succeed if you say it is confusing. Could we 
revive the paragraph and explain better? I can give it another try!}

{\john  Okay, that makes sense.   I put the paragraph back in and changed it a bunch: basically, I added some more steps so people who don't know what 
we're doing will have an easier time figuring it out.   Does it look okay to you?}

{\chris Yes it does, thanks! Although in my mind, it kinda goes the opposite way: e.g. the already induced monoidal product on the fibers ends up 
having the universal property of a coproduct for example. Possibly just overthinking and getting mixed up with words.}

{\john I changed the paragraph, trying to solve the problem Christina just mentioned.}

@@

{\john $\checkmark$} 
\iffalse
{\kenny I think the reader should come to their own conclusion about which framework they should use if they are trying to decide between the two, assuming both are applicable. I wouldn't say that one is better than the other. Personally, I would rather try structured cospans first, because I would rather think about a left adjoint than a symmetric lax monoidal pseudofunctor, but other people might feel differently.}

{\john I agree with you, Kenny.  I think this reviewer saw the word ``vs'' and thought we were going to let structured and decorated cospans fight it out and announce who won, like at a boxing match.   Instead of arguing that this is not a good idea for a paper, I decided to just point out that our abstract says what our paper does, and it's something else.   (I also agree with you that structured cospans are simpler when you can use them: that's why we invented them!)}

{\john If you like the response in boldface above, please add a checkmark here.  Otherwise try to
improve it, or something.}  
\fi

{\chris I could add ``Those examples should serve as a guide for other similar applications, 
whereas new applications would require exploring which approach works best, for example if a left adjoint (to start off with structured cospans) or a 
lax monoidal pseudofunctor (to start off with decorated cospans) is available. We do not make a stand that either is strictly better.'' In fact, we 
could add something soft along those lines somewhere in the introduction to satisfy the reviewer, if we like.}

@@

{\chris Sounds like a great idea to me!! I was also a bit scared to go into this when I first saw it, but I kinda thought morally we should, and now 
it looks like it won't be a disaster, and in fact a very fruitful result!! So yes please, do it and let me know what I can do to help except 
proofread it.}

{\chris Also perhaps we should add those references they provide somewhere.}

\item If we take the categorical systems theory perspective advanced by section 6, and the body of the authors’ recent work on functorial semantics 
for complex 
systems, then we should ask why the decorated and structured cospan double categories are isomorphic for reaction networks, but their semantics as 
dynamical systems are a decorated cospan category that is not equivalent to any structured cospan category. Is there a different construction of 
semantics for reaction networks that uses structured cospans?

{\bf We do not know a way to construct a (double) category of open dynamical systems that uses structured cospans; to find one would require getting around the result at the end of Section 6.  This is an interesting challenge.}

{\john $\checkmark$}

@@

{\chris Should we refer to which result we are talking about? Also, I am a bit confused about the question and answer. I thought reaction networks 
can be described in both ways?}

{\kenny I'm kinda' confused by the question, too, due to its wording. Based on what Dr. Baez said in response, I think they might have meant dynamical systems instead of reaction networks in their last sentence. They even say earlier in their comment that the frameworks are isomorphic for reaction networks.}

{\john The referee was asking about the \emph{semantics} for reaction networks --- that is, dynamical systems!  Reaction networks are the ``syntax'' and dynamical systems are our chosen ``semantics'' for these.  But my reply in boldface was confused --- you're right.  So I've fixed it.  Give me a check mark if you're happy now.}

### 2022-03-15 cospans_new.tex
dealt with comment 11.

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we apparently need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.

@@

The hypothesis of this theorem is essentially a way of asking that the fibers of the induced opfibration $U \maps \inta F \to \A$ have finite colimits which are preserved by the reindexing functors, and that the induced fiberwise monoidal structure is cocartesian.   Expanding on this a bit:
the fibers of $U$ are the same as the categories $F(a)$ for all $a \in \A$.   The lax monoidal pseudofunctor structure of $F$ gives rise to a specific symmetric monoidal structure on the categories $F(a)$ given by \cref{eq:explicitstructure1}, and the hypothesis of the theorem asks that the resulting pseudofunctor $F \maps \A \to \SMC$ factors through $\Rex$.  This implies
that each category $F(a)$ has coproducts and an initial object, and that the resulting cocartesian structure is what makes $F(a)$ a symmetric monoidal category.

>>>>

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we apparently need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.

@@

The hypothesis of this theorem is essentially a way of asking that the fibers of the induced opfibration $U \maps \inta F \to \A$ have finite colimits which are preserved by the reindexing functors, and that the induced fiberwise monoidal structure is cocartesian.   Expanding on this a bit:
the fibers of $U$ are the same as the categories $F(a)$ for all $a \in \A$.   The lax monoidal pseudofunctor structure of $F$ gives rise to a specific symmetric monoidal structure on each category $F(a)$ given by \cref{eq:explicitstructure1}.  The hypothesis of the theorem asks that the resulting pseudofunctor $F \maps \A \to \SMC$ factors through $\Rex$.  This implies
that the symmetric monoidal structure on $F(a)$ is cocartesian.
